---
title: "230823_Covid_NSAID_APAP_workflow"
author: "Aurélie Pahud de Mortanges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(tidyverse)
library(dplyr)
library(writexl)
library(caret)
library(magrittr)
library(rpart)
library(rpart.plot)
library(gridExtra)
library(dendroTools)
library(randomForest)
library(imputeTS)
library(zoo)

```


# CASE CONTROL STUDY NSAID/KIDNEY

## Modeling

### case control study NSAID/kidney: prepare for modeling 
```{r}

load(file="data_kidney.Rdata")

# split train and test set
set.seed(42)
index <- createDataPartition(y = ren_plaus_dem_to_model$outcome, p = 0.8, list = F, times = 1)

train_data_kidney <- ren_plaus_dem_to_model[index,]
test_data_kidney <- ren_plaus_dem_to_model[-index,]

# standardize data
pp <- preProcess(train_data_kidney, method = "range", rangeBounds = c(0,1))
train_data_kidney_stand <- predict(pp, train_data_kidney)
test_data_kidney_stand <- predict(pp, test_data_kidney)

# set up model weights
my_weights <- if_else(train_data_kidney$outcome == "case",
                      (1/sum(train_data_kidney$outcome == "case")) * 0.5,
                      (1/sum(train_data_kidney$outcome == "control")) * 0.5)

```



### case control study NSAID/kidney: logistic regression with caret
```{r warning=FALSE}

# set up train control
lr_train_control <- trainControl(method = "cv",
                                 number = 5,
                                 classProbs = TRUE,
                                 summaryFunction = twoClassSummary,
                                 seeds = c(1:6))

set.seed(42)
lr_model <- caret::train(outcome ~ ., 
                         data = train_data_kidney_stand,
                         trControl = lr_train_control,
                         weights = my_weights,
                         method = "glm",
                         family = "binomial",
                         metric = "ROC")

# calculate confusion matrix and metrics train
lr_predicted_values_train <- predict(lr_model, newdata = train_data_kidney_stand)
lr_actual_values_train <- as.factor(train_data_kidney$outcome)
lr_conf_matrix_train <- confusionMatrix(lr_predicted_values_train, lr_actual_values_train, positive = "case")
lr_conf_matrix_train

# calculate confusion matrix test
lr_predicted_values_test <- predict(lr_model, newdata = test_data_kidney_stand)
lr_actual_values_test <- as.factor(test_data_kidney_stand$outcome)
lr_conf_matrix_test <- confusionMatrix(lr_predicted_values_test, lr_actual_values_test, positive = "case")
lr_conf_matrix_test

rm(lr_predicted_values_train, lr_actual_values_train, lr_predicted_values_test, lr_actual_values_test)

```

### case control study NSAID/kidney: custom tuned decision tree
```{r}

set.seed(42)
dt_model2 <- rpart(outcome ~ ., 
                   data = train_data_kidney, 
                   method = "class",
                   cp = 0.002, minbucket = 9, minsplit = 5, maxdepth = 8)

dt_predicted_values_train2 <- predict(dt_model2, newdata = train_data_kidney, type = "class")
dt_actual_values_train2 <- as.factor(train_data_kidney$outcome)
dt_conf_matrix_train2 <- confusionMatrix(dt_predicted_values_train2, dt_actual_values_train2, positive = "case")
dt_conf_matrix_train2

dt_predicted_values_test2 <- predict(dt_model2, newdata = test_data_kidney, type = "class")
dt_actual_values_test2 <- as.factor(test_data_kidney$outcome)
dt_conf_matrix_test2 <- confusionMatrix(dt_predicted_values_test2, dt_actual_values_test2, positive = "case")
dt_conf_matrix_test2

# make nice visualization
pdf("decisionTree_NSAID_231101.pdf", height = 5, width = 7)
prp(dt_model2, branch = 1, box.palette = c("#a4c3d8", "#eb8c5a"), cex = 0.5, type = 5, varlen = 12)
dev.off()

rm(dt_predicted_values_train2, dt_actual_values_train2, dt_predicted_values_test2, dt_actual_values_test2)

```

### case control study NSAID/kidney: knn with caret
```{r}

# set up train control
knn_train_control <- trainControl(method = "cv",
                                 number = 5,
                                 classProbs = TRUE,
                                 summaryFunction = twoClassSummary,
                                 seeds = list(c(1:10), c(11:20), c(21:30), c(31:40), c(41:50), 51))

# knn needs standardized data
set.seed(42)
knn_model <- caret::train(outcome ~., data = train_data_kidney_stand, 
                   method = "knn", 
                   trControl = knn_train_control, 
                   tuneLength = 10,
                   weights = my_weights,
                   metric = "ROC") 

# calculate confusion matrix and metrics train
knn_predicted_values_train <- predict(knn_model, newdata = train_data_kidney_stand)
knn_actual_values_train <- as.factor(train_data_kidney$outcome)
knn_conf_matrix_train <- confusionMatrix(knn_predicted_values_train, knn_actual_values_train, positive = "case")
knn_conf_matrix_train

# calculate confusion matrix test
knn_predicted_values_test <- predict(knn_model, newdata = test_data_kidney_stand)
knn_actual_values_test <- as.factor(test_data_kidney_stand$outcome)
knn_conf_matrix_test <- confusionMatrix(knn_predicted_values_test, knn_actual_values_test, positive = "case")
knn_conf_matrix_test

rm(knn_predicted_values_train, knn_actual_values_train, knn_predicted_values_test, knn_actual_values_test)
```


### case control study NSAID/kidney: custom tuned random Forest from Verena, adapted, part 2
```{r}

# run RF with optimal parameters
set.seed(42)
rf_model2 <- randomForest(x = train_data_kidney[, -4], y = as.factor(train_data_kidney$outcome) , maxnodes = 5, ntree = 60, weights = my_weights)

# calculate confusion matrix train
rf_predicted_values_train2 <- predict(rf_model2, train_data_kidney[, -4])
rf_actual_values_train2 <- as.factor(train_data_kidney$outcome)
rf_conf_matrix_train2 <- confusionMatrix(rf_predicted_values_train2, rf_actual_values_train2, positive = "case")
rf_conf_matrix_train2

# calculate confusion matrix test
rf_predicted_values_test2 <- predict(rf_model2, test_data_kidney[, -4])
rf_actual_values_test2 <- as.factor(test_data_kidney$outcome)
rf_conf_matrix_test2 <- confusionMatrix(rf_predicted_values_test2, rf_actual_values_test2, positive = "case")
rf_conf_matrix_test2

rm(rf_predicted_values_train2, rf_predicted_values_test2, rf_actual_values_train2, rf_actual_values_test2)
```

### case control study NSAID/kidney: AdaBoost with caret
```{r}

# set up train control
ada_train_control <- trainControl(method = "cv",
                                 number = 5,
                                 classProbs = TRUE,
                                 summaryFunction = twoClassSummary)

ada_grid <- expand.grid(mfinal = seq(100, 200, by = 20), 
                       maxdepth = seq(2, 5, by = 1),
                       coeflearn = c("Breiman", "Freund"))

set.seed(42)
ada_model <- caret::train(outcome ~ ., 
                  data = train_data_kidney, 
                  method = "AdaBoost.M1", 
                  tuneGrid = ada_grid,
                  trControl = ada_train_control,
                  weights = my_weights,
                  metric = "ROC")

# calculate confusion matrix train
ada_predicted_values_train <- predict(ada_model, newdata = train_data_kidney)
ada_actual_values_train <- as.factor(train_data_kidney$outcome)
ada_conf_matrix_train <- confusionMatrix(ada_predicted_values_train, ada_actual_values_train, positive = "case")
ada_conf_matrix_train

# calculate confusion matrix test
ada_predicted_values_test <- predict(ada_model, newdata = test_data_kidney)
ada_actual_values_test <- as.factor(test_data_kidney$outcome)
ada_conf_matrix_test <- confusionMatrix(ada_predicted_values_test, ada_actual_values_test, positive = "case")
ada_conf_matrix_test

rm(ada_predicted_values_train, ada_predicted_values_test, ada_actual_values_train, ada_actual_values_test)

```

## Results

###  case control study NSAID/kidney: assemble results
```{r}

# logistic regression caret
lr_overall_train <- lr_conf_matrix_train$overall
lr_accConf_train <- paste(round(lr_overall_train[["Accuracy"]], digits = 2), " (", round(lr_overall_train[["AccuracyLower"]], digits = 2), " - ", round(lr_overall_train[["AccuracyUpper"]], digits = 2), ")", sep = "")
lr_byClass_train <- lr_conf_matrix_train$byClass
lr_sens_train <- lr_byClass_train["Sensitivity"]
lr_spec_train <- lr_byClass_train["Specificity"]
lr_PPV_train <- lr_byClass_train["Pos Pred Value"]
lr_NPV_train <- lr_byClass_train["Neg Pred Value"]
lr_prec_train <- lr_byClass_train["Precision"]
lr_rec_train <- lr_byClass_train["Recall"]
lr_F1_train <- lr_byClass_train["F1"]
lr_balAc_train <- lr_byClass_train["Balanced Accuracy"]

lr_overall_test <- lr_conf_matrix_test$overall
lr_accConf_test <- paste(round(lr_overall_test[["Accuracy"]], digits = 2), " (", round(lr_overall_test[["AccuracyLower"]], digits = 2), " - ", round(lr_overall_test[["AccuracyUpper"]], digits = 2), ")", sep = "")
lr_byClass_test <- lr_conf_matrix_test$byClass
lr_sens_test <- lr_byClass_test["Sensitivity"]
lr_spec_test <- lr_byClass_test["Specificity"]
lr_PPV_test <- lr_byClass_test["Pos Pred Value"]
lr_NPV_test <- lr_byClass_test["Neg Pred Value"]
lr_prec_test <- lr_byClass_test["Precision"]
lr_rec_test <- lr_byClass_test["Recall"]
lr_F1_test <- lr_byClass_test["F1"]
lr_balAc_test <- lr_byClass_test["Balanced Accuracy"]

LR_train = c(lr_balAc_train, lr_sens_train, lr_spec_train, lr_PPV_train, lr_NPV_train, lr_prec_train, lr_rec_train, lr_F1_train)
LR_test = c(lr_balAc_test, lr_sens_test, lr_spec_test, lr_PPV_test, lr_NPV_test, lr_prec_test, lr_rec_test, lr_F1_test)

rm(lr_overall_train, lr_byClass_train, lr_sens_train, lr_spec_train, lr_PPV_train, lr_NPV_train, lr_prec_train, lr_rec_train, lr_F1_train, lr_balAc_train, lr_overall_test, lr_byClass_test, lr_sens_test, lr_spec_test, lr_PPV_test, lr_NPV_test, lr_prec_test, lr_rec_test, lr_F1_test, lr_balAc_test)

# decision tree custom
dt2_overall_train <- dt_conf_matrix_train2$overall
dt2_accConf_train <- paste(round(dt2_overall_train[["Accuracy"]], digits = 2), " (", round(dt2_overall_train[["AccuracyLower"]], digits = 2), " - ", round(dt2_overall_train[["AccuracyUpper"]], digits = 2), ")", sep = "")
dt2_byClass_train <- dt_conf_matrix_train2$byClass
dt2_sens_train <- dt2_byClass_train["Sensitivity"]
dt2_spec_train <- dt2_byClass_train["Specificity"]
dt2_PPV_train <- dt2_byClass_train["Pos Pred Value"]
dt2_NPV_train <- dt2_byClass_train["Neg Pred Value"]
dt2_prec_train <- dt2_byClass_train["Precision"]
dt2_rec_train <- dt2_byClass_train["Recall"]
dt2_F1_train <- dt2_byClass_train["F1"]
dt2_balAc_train <- dt2_byClass_train["Balanced Accuracy"]

dt2_overall_test <- dt_conf_matrix_test2$overall
dt2_accConf_test <- paste(round(dt2_overall_test[["Accuracy"]], digits = 2), " (", round(dt2_overall_test[["AccuracyLower"]], digits = 2), " - ", round(dt2_overall_test[["AccuracyUpper"]], digits = 2), ")", sep = "")
dt2_byClass_test <- dt_conf_matrix_test2$byClass
dt2_sens_test <- dt2_byClass_test["Sensitivity"]
dt2_spec_test <- dt2_byClass_test["Specificity"]
dt2_PPV_test <- dt2_byClass_test["Pos Pred Value"]
dt2_NPV_test <- dt2_byClass_test["Neg Pred Value"]
dt2_prec_test <- dt2_byClass_test["Precision"]
dt2_rec_test <- dt2_byClass_test["Recall"]
dt2_F1_test <- dt2_byClass_test["F1"]
dt2_balAc_test <- dt2_byClass_test["Balanced Accuracy"]

DT_train = c(dt2_balAc_train, dt2_sens_train, dt2_spec_train, dt2_PPV_train, dt2_NPV_train, dt2_prec_train, dt2_rec_train, dt2_F1_train)
DT_test = c(dt2_balAc_test, dt2_sens_test, dt2_spec_test, dt2_PPV_test, dt2_NPV_test, dt2_prec_test, dt2_rec_test, dt2_F1_test)

rm(dt2_sens_train, dt2_spec_train, dt2_PPV_train, dt2_NPV_train, dt2_prec_train, dt2_rec_train, dt2_F1_train, dt2_balAc_train, dt2_sens_test, dt2_spec_test, dt2_PPV_test, dt2_NPV_test, dt2_prec_test, dt2_rec_test, dt2_F1_test, dt2_balAc_test)

# knn caret
knn_overall_train <- knn_conf_matrix_train$overall
knn_accConf_train <- paste(round(knn_overall_train[["Accuracy"]], digits = 2), " (", round(knn_overall_train[["AccuracyLower"]], digits = 2), " - ", round(knn_overall_train[["AccuracyUpper"]], digits = 2), ")", sep = "")
knn_byClass_train <- knn_conf_matrix_train$byClass
knn_sens_train <- knn_byClass_train["Sensitivity"]
knn_spec_train <- knn_byClass_train["Specificity"]
knn_PPV_train <- knn_byClass_train["Pos Pred Value"]
knn_NPV_train <- knn_byClass_train["Neg Pred Value"]
knn_prec_train <- knn_byClass_train["Precision"]
knn_rec_train <- knn_byClass_train["Recall"]
knn_F1_train <- knn_byClass_train["F1"]
knn_balAc_train <- knn_byClass_train["Balanced Accuracy"]

knn_overall_test <- knn_conf_matrix_test$overall
knn_accConf_test <- paste(round(knn_overall_test[["Accuracy"]], digits = 2), " (", round(knn_overall_test[["AccuracyLower"]], digits = 2), " - ", round(knn_overall_test[["AccuracyUpper"]], digits = 2), ")", sep = "")
knn_byClass_test <- knn_conf_matrix_test$byClass
knn_sens_test <- knn_byClass_test["Sensitivity"]
knn_spec_test <- knn_byClass_test["Specificity"]
knn_PPV_test <- knn_byClass_test["Pos Pred Value"]
knn_NPV_test <- knn_byClass_test["Neg Pred Value"]
knn_prec_test <- knn_byClass_test["Precision"]
knn_rec_test <- knn_byClass_test["Recall"]
knn_F1_test <- knn_byClass_test["F1"]
knn_balAc_test <- knn_byClass_test["Balanced Accuracy"]

KNN_train = c(knn_balAc_train, knn_sens_train, knn_spec_train, knn_PPV_train, knn_NPV_train, knn_prec_train, knn_rec_train, knn_F1_train)
KNN_test = c(knn_balAc_test, knn_sens_test, knn_spec_test, knn_PPV_test, knn_NPV_test, knn_prec_test, knn_rec_test, knn_F1_test)

rm(knn_sens_train, knn_spec_train, knn_PPV_train, knn_NPV_train, knn_prec_train, knn_rec_train, knn_F1_train, knn_balAc_train, knn_sens_test, knn_spec_test, knn_PPV_test, knn_NPV_test, knn_prec_test, knn_rec_test, knn_F1_test, knn_balAc_test)

# random forest custom
rf2_overall_train <- rf_conf_matrix_train2$overall
rf2_accConf_train <- paste(round(rf2_overall_train[["Accuracy"]], digits = 2), " (", round(rf2_overall_train[["AccuracyLower"]], digits = 2), " - ", round(rf2_overall_train[["AccuracyUpper"]], digits = 2), ")", sep = "")
rf2_byClass_train <- rf_conf_matrix_train2$byClass
rf2_sens_train <- rf2_byClass_train["Sensitivity"]
rf2_spec_train <- rf2_byClass_train["Specificity"]
rf2_PPV_train <- rf2_byClass_train["Pos Pred Value"]
rf2_NPV_train <- rf2_byClass_train["Neg Pred Value"]
rf2_prec_train <- rf2_byClass_train["Precision"]
rf2_rec_train <- rf2_byClass_train["Recall"]
rf2_F1_train <- rf2_byClass_train["F1"]
rf2_balAc_train <- rf2_byClass_train["Balanced Accuracy"]

rf2_overall_test <- rf_conf_matrix_test2$overall
rf2_accConf_test <- paste(round(rf2_overall_test[["Accuracy"]], digits = 2), " (", round(rf2_overall_test[["AccuracyLower"]], digits = 2), " - ", round(rf2_overall_test[["AccuracyUpper"]], digits = 2), ")", sep = "")
rf2_byClass_test <- rf_conf_matrix_test2$byClass
rf2_sens_test <- rf2_byClass_test["Sensitivity"]
rf2_spec_test <- rf2_byClass_test["Specificity"]
rf2_PPV_test <- rf2_byClass_test["Pos Pred Value"]
rf2_NPV_test <- rf2_byClass_test["Neg Pred Value"]
rf2_prec_test <- rf2_byClass_test["Precision"]
rf2_rec_test <- rf2_byClass_test["Recall"]
rf2_F1_test <- rf2_byClass_test["F1"]
rf2_balAc_test <- rf2_byClass_test["Balanced Accuracy"]

RF_train = c(rf2_balAc_train, rf2_sens_train, rf2_spec_train, rf2_PPV_train, rf2_NPV_train, rf2_prec_train, rf2_rec_train, rf2_F1_train)
RF_test = c(rf2_balAc_test, rf2_sens_test, rf2_spec_test, rf2_PPV_test, rf2_NPV_test, rf2_prec_test, rf2_rec_test, rf2_F1_test)

rm(rf2_sens_train, rf2_spec_train, rf2_PPV_train, rf2_NPV_train, rf2_prec_train, rf2_rec_train, rf2_F1_train, rf2_balAc_train, rf2_sens_test, rf2_spec_test, rf2_PPV_test, rf2_NPV_test, rf2_prec_test, rf2_rec_test, rf2_F1_test, rf2_balAc_test)

# AdaBoost caret
ada_overall_train <- ada_conf_matrix_train$overall
ada_accConf_train <- paste(round(ada_overall_train[["Accuracy"]], digits = 2), " (", round(ada_overall_train[["AccuracyLower"]], digits = 2), " - ", round(ada_overall_train[["AccuracyUpper"]], digits = 2), ")", sep = "")
ada_byClass_train <- ada_conf_matrix_train$byClass
ada_sens_train <- ada_byClass_train["Sensitivity"]
ada_spec_train <- ada_byClass_train["Specificity"]
ada_PPV_train <- ada_byClass_train["Pos Pred Value"]
ada_NPV_train <- ada_byClass_train["Neg Pred Value"]
ada_prec_train <- ada_byClass_train["Precision"]
ada_rec_train <- ada_byClass_train["Recall"]
ada_F1_train <- ada_byClass_train["F1"]
ada_balAc_train <- ada_byClass_train["Balanced Accuracy"]

ada_overall_test <- ada_conf_matrix_test$overall
ada_accConf_test <- paste(round(ada_overall_test[["Accuracy"]], digits = 2), " (", round(ada_overall_test[["AccuracyLower"]], digits = 2), " - ", round(ada_overall_test[["AccuracyUpper"]], digits = 2), ")", sep = "")
ada_byClass_test <- ada_conf_matrix_test$byClass
ada_sens_test <- ada_byClass_test["Sensitivity"]
ada_spec_test <- ada_byClass_test["Specificity"]
ada_PPV_test <- ada_byClass_test["Pos Pred Value"]
ada_NPV_test <- ada_byClass_test["Neg Pred Value"]
ada_prec_test <- ada_byClass_test["Precision"]
ada_rec_test <- ada_byClass_test["Recall"]
ada_F1_test <- ada_byClass_test["F1"]
ada_balAc_test <- ada_byClass_test["Balanced Accuracy"]

AB_train = c(ada_balAc_train, ada_sens_train, ada_spec_train, ada_PPV_train, ada_NPV_train, ada_prec_train, ada_rec_train, ada_F1_train)
AB_test = c(ada_balAc_test, ada_sens_test, ada_spec_test, ada_PPV_test, ada_NPV_test, ada_prec_test, ada_rec_test, ada_F1_test)

rm(ada_overall_train, ada_sens_train, ada_spec_train, ada_PPV_train, ada_NPV_train, ada_prec_train, ada_rec_train, ada_F1_train, ada_balAc_train, ada_overall_test, ada_sens_test, ada_spec_test, ada_PPV_test, ada_NPV_test, ada_prec_test, ada_rec_test, ada_F1_test, ada_balAc_test)

# assemble table
results_NSAID_table <- round_df(data.frame(LR_train, LR_test, DT_train, DT_test, RF_train, RF_test, KNN_train, KNN_test, AB_train, AB_test), digits = 2)

# accuracy
accConf <- c(lr_accConf_train, lr_accConf_test, dt2_accConf_train, dt2_accConf_test, rf2_accConf_train, rf2_accConf_test, knn_accConf_train, knn_accConf_test, ada_accConf_train, ada_accConf_test)

results_NSAID_table <- rbind(accConf, results_NSAID_table)
rownames(results_NSAID_table) <- c("Accuracy", "Balanced Accuracy", "Sensitivity", "Specificity", "PPV", "NPV", "Precision", "Recall", "F1 Score")

write_xlsx(results_NSAID_table, "results_NSAIDs_231101.xlsx")

pdf("results_NSAIDs_231109.pdf", width = 18, height = 5)
results_NSAID_table <- tableGrob(data.frame(results_NSAID_table))
grid.arrange(results_NSAID_table)
dev.off()

```

# CASE CONTROL STUDY APAP/LIVER

## Modeling

### case control study APAP/liver: prepare for modeling 
```{r}

load(file="data_liver.Rdata")

# split train and test set
set.seed(42)
index <- createDataPartition(y = hep_plaus_dem_to_model$outcome, p = 0.8, list = F, times = 1)

train_data_liver <- hep_plaus_dem_to_model[index,]
test_data_liver <- hep_plaus_dem_to_model[-index,]

# standardize data
pp <- preProcess(train_data_liver, method = "range", rangeBounds = c(0,1))
train_data_liver_stand <- predict(pp, train_data_liver)
test_data_liver_stand <- predict(pp, test_data_liver)

# set up train control
my_train_control <- trainControl(method = "cv",
                                 number = 5,
                                 classProbs = TRUE,
                                 summaryFunction = twoClassSummary)

# set up model weights
my_weights <- if_else(train_data_liver$outcome == "case",
                      (1/sum(train_data_liver$outcome == "case")) * 0.5,
                      (1/sum(train_data_liver$outcome == "control")) * 0.5)

```


### case control study APAP/liver: logistic regression with caret
```{r warning=FALSE}
set.seed(42)

lr_model <- caret::train(outcome ~ ., 
                         data = train_data_liver_stand,
                         trControl = my_train_control,
                         weights = my_weights,
                         method = "glm",
                         family = "binomial",
                         metric = "ROC")

# calculate confusion matrix and metrics train
lr_predicted_values_train <- predict(lr_model, newdata = train_data_liver_stand)
lr_actual_values_train <- as.factor(train_data_liver$outcome)
lr_conf_matrix_train <- confusionMatrix(lr_predicted_values_train, lr_actual_values_train, positive = "case")
lr_conf_matrix_train

# calculate confusion matrix test
lr_predicted_values_test <- predict(lr_model, newdata = test_data_liver_stand)
lr_actual_values_test <- as.factor(test_data_liver_stand$outcome)
lr_conf_matrix_test <- confusionMatrix(lr_predicted_values_test, lr_actual_values_test, positive = "case")
lr_conf_matrix_test

rm(lr_predicted_values_train, lr_actual_values_train, lr_predicted_values_test, lr_actual_values_test)

```

### case control study APAP/liver: custom tuned decision tree
```{r}

set.seed(42)
dt_model2 <- rpart(outcome ~ ., 
                   data = train_data_liver, 
                   method = "class",
                   cp = 0.001, minbucket = 8, minsplit = 3, maxdepth = 5)

dt_predicted_values_train2 <- predict(dt_model2, newdata = train_data_liver, type = "class")
dt_actual_values_train2 <- as.factor(train_data_liver$outcome)
dt_conf_matrix_train2 <- confusionMatrix(dt_predicted_values_train2, dt_actual_values_train2, positive = "case")
dt_conf_matrix_train2

dt_predicted_values_test2 <- predict(dt_model2, newdata = test_data_liver, type = "class")
dt_actual_values_test2 <- as.factor(test_data_liver$outcome)
dt_conf_matrix_test2 <- confusionMatrix(dt_predicted_values_test2, dt_actual_values_test2, positive = "case")
dt_conf_matrix_test2

# make nice visualization
pdf("decisionTree_APAP_231109.pdf", height = 5, width = 7)
prp(dt_model2, branch = 1, box.palette = c("#a4c3d8", "#eb8c5a"), cex = 0.5, type = 5, varlen = 12)
dev.off()

rm(dt_predicted_values_train2, dt_actual_values_train2, dt_predicted_values_test2, dt_actual_values_test2)

```

### case control study APAP/liver: knn with caret
```{r}
set.seed(2)

knn_model <- caret::train(outcome ~., data = train_data_liver_stand,
                          method = "knn", 
                          trControl = my_train_control, 
                          tuneLength = 10,
                          weights = my_weights,
                          metric = "ROC") 

# calculate confusion matrix and metrics train
knn_predicted_values_train <- predict(knn_model, newdata = train_data_liver_stand)
knn_actual_values_train <- as.factor(train_data_liver$outcome)
knn_conf_matrix_train <- confusionMatrix(knn_predicted_values_train, knn_actual_values_train, positive = "case")
knn_conf_matrix_train

# calculate confusion matrix test
knn_predicted_values_test <- predict(knn_model, newdata = test_data_liver_stand)
knn_actual_values_test <- as.factor(test_data_liver_stand$outcome)
knn_conf_matrix_test <- confusionMatrix(knn_predicted_values_test, knn_actual_values_test, positive = "case")
knn_conf_matrix_test

rm(knn_predicted_values_train, knn_actual_values_train, knn_predicted_values_test, knn_actual_values_test)
```

### case control study APAP/livery: custom tuned random Forest 
```{r}
set.seed(42)

# run RF with optimal parameters
rf_model2 <- randomForest(x = train_data_liver[, -4], y = train_data_liver[, 4] , maxnodes = 8, ntree = 70, weights = my_weights)
rf_model2

# calculate confusion matrix train
rf_predicted_values_train2 <- predict(rf_model2, train_data_liver[, -4])
rf_actual_values_train2 <- as.factor(train_data_liver[, 4])
rf_conf_matrix_train2 <- confusionMatrix(rf_predicted_values_train2, rf_actual_values_train2, positive = "case")
rf_conf_matrix_train2

# calculate confusion matrix test
rf_predicted_values_test2 <- predict(rf_model2, test_data_liver[, -4])
rf_actual_values_test2 <- as.factor(test_data_liver[, 4])
rf_conf_matrix_test2 <- confusionMatrix(rf_predicted_values_test2, rf_actual_values_test2, positive = "case")
rf_conf_matrix_test2

rm(rf_predicted_values_train2, rf_predicted_values_test2, rf_actual_values_train2, rf_actual_values_test2)
```

### case control study APAP/liver: AdaBoost with caret
```{r}

set.seed(42)

fitGrid <- expand.grid(mfinal = seq(100, 200, by = 20), 
                       maxdepth = seq(2, 5, by = 1),
                       coeflearn = c("Breiman", "Freund"))

ada_model <- caret::train(outcome ~ ., 
                  data = train_data_liver, 
                  method = "AdaBoost.M1", 
                  tuneGrid = fitGrid,
                  trControl = my_train_control,
                  weights = my_weights,
                  metric = "ROC")

# calculate confusion matrix train
ada_predicted_values_train <- predict(ada_model, newdata = train_data_liver)
ada_actual_values_train <- as.factor(train_data_liver$outcome)
ada_conf_matrix_train <- confusionMatrix(ada_predicted_values_train, ada_actual_values_train, positive = "case")
ada_conf_matrix_train

# calculate confusion matrix test
ada_predicted_values_test <- predict(ada_model, newdata = test_data_liver)
ada_actual_values_test <- as.factor(test_data_liver$outcome)
ada_conf_matrix_test <- confusionMatrix(ada_predicted_values_test, ada_actual_values_test, positive = "case")
ada_conf_matrix_test

rm(ada_predicted_values_test, ada_predicted_values_train, ada_actual_values_test, ada_actual_values_train)

```

## Results

### case control study APAP/liver: assemble results
```{r}

# logistic regression caret
lr_overall_train <- lr_conf_matrix_train$overall
lr_accConf_train <- paste(round(lr_overall_train[["Accuracy"]], digits = 2), " (", round(lr_overall_train[["AccuracyLower"]], digits = 2), " - ", round(lr_overall_train[["AccuracyUpper"]], digits = 2), ")", sep = "")
lr_byClass_train <- lr_conf_matrix_train$byClass
lr_sens_train <- lr_byClass_train["Sensitivity"]
lr_spec_train <- lr_byClass_train["Specificity"]
lr_PPV_train <- lr_byClass_train["Pos Pred Value"]
lr_NPV_train <- lr_byClass_train["Neg Pred Value"]
lr_prec_train <- lr_byClass_train["Precision"]
lr_rec_train <- lr_byClass_train["Recall"]
lr_F1_train <- lr_byClass_train["F1"]
lr_balAc_train <- lr_byClass_train["Balanced Accuracy"]

lr_overall_test <- lr_conf_matrix_test$overall
lr_accConf_test <- paste(round(lr_overall_test[["Accuracy"]], digits = 2), " (", round(lr_overall_test[["AccuracyLower"]], digits = 2), " - ", round(lr_overall_test[["AccuracyUpper"]], digits = 2), ")", sep = "")
lr_byClass_test <- lr_conf_matrix_test$byClass
lr_sens_test <- lr_byClass_test["Sensitivity"]
lr_spec_test <- lr_byClass_test["Specificity"]
lr_PPV_test <- lr_byClass_test["Pos Pred Value"]
lr_NPV_test <- lr_byClass_test["Neg Pred Value"]
lr_prec_test <- lr_byClass_test["Precision"]
lr_rec_test <- lr_byClass_test["Recall"]
lr_F1_test <- lr_byClass_test["F1"]
lr_balAc_test <- lr_byClass_test["Balanced Accuracy"]

LR_train = c(lr_balAc_train, lr_sens_train, lr_spec_train, lr_PPV_train, lr_NPV_train, lr_prec_train, lr_rec_train, lr_F1_train)
LR_test = c(lr_balAc_test, lr_sens_test, lr_spec_test, lr_PPV_test, lr_NPV_test, lr_prec_test, lr_rec_test, lr_F1_test)

rm(lr_overall_train, lr_byClass_train, lr_sens_train, lr_spec_train, lr_PPV_train, lr_NPV_train, lr_prec_train, lr_rec_train, lr_F1_train, lr_balAc_train, lr_overall_test, lr_byClass_test, lr_sens_test, lr_spec_test, lr_PPV_test, lr_NPV_test, lr_prec_test, lr_rec_test, lr_F1_test, lr_balAc_test)

# decision tree custom
dt2_overall_train <- dt_conf_matrix_train2$overall
dt2_accConf_train <- paste(round(dt2_overall_train[["Accuracy"]], digits = 2), " (", round(dt2_overall_train[["AccuracyLower"]], digits = 2), " - ", round(dt2_overall_train[["AccuracyUpper"]], digits = 2), ")", sep = "")
dt2_byClass_train <- dt_conf_matrix_train2$byClass
dt2_sens_train <- dt2_byClass_train["Sensitivity"]
dt2_spec_train <- dt2_byClass_train["Specificity"]
dt2_PPV_train <- dt2_byClass_train["Pos Pred Value"]
dt2_NPV_train <- dt2_byClass_train["Neg Pred Value"]
dt2_prec_train <- dt2_byClass_train["Precision"]
dt2_rec_train <- dt2_byClass_train["Recall"]
dt2_F1_train <- dt2_byClass_train["F1"]
dt2_balAc_train <- dt2_byClass_train["Balanced Accuracy"]

dt2_overall_test <- dt_conf_matrix_test2$overall
dt2_accConf_test <- paste(round(dt2_overall_test[["Accuracy"]], digits = 2), " (", round(dt2_overall_test[["AccuracyLower"]], digits = 2), " - ", round(dt2_overall_test[["AccuracyUpper"]], digits = 2), ")", sep = "")
dt2_byClass_test <- dt_conf_matrix_test2$byClass
dt2_sens_test <- dt2_byClass_test["Sensitivity"]
dt2_spec_test <- dt2_byClass_test["Specificity"]
dt2_PPV_test <- dt2_byClass_test["Pos Pred Value"]
dt2_NPV_test <- dt2_byClass_test["Neg Pred Value"]
dt2_prec_test <- dt2_byClass_test["Precision"]
dt2_rec_test <- dt2_byClass_test["Recall"]
dt2_F1_test <- dt2_byClass_test["F1"]
dt2_balAc_test <- dt2_byClass_test["Balanced Accuracy"]

DT_train = c(dt2_balAc_train, dt2_sens_train, dt2_spec_train, dt2_PPV_train, dt2_NPV_train, dt2_prec_train, dt2_rec_train, dt2_F1_train)
DT_test = c(dt2_balAc_test, dt2_sens_test, dt2_spec_test, dt2_PPV_test, dt2_NPV_test, dt2_prec_test, dt2_rec_test, dt2_F1_test)

rm(dt2_sens_train, dt2_spec_train, dt2_PPV_train, dt2_NPV_train, dt2_prec_train, dt2_rec_train, dt2_F1_train, dt2_balAc_train, dt2_sens_test, dt2_spec_test, dt2_PPV_test, dt2_NPV_test, dt2_prec_test, dt2_rec_test, dt2_F1_test, dt2_balAc_test)

# knn caret
knn_overall_train <- knn_conf_matrix_train$overall
knn_accConf_train <- paste(round(knn_overall_train[["Accuracy"]], digits = 2), " (", round(knn_overall_train[["AccuracyLower"]], digits = 2), " - ", round(knn_overall_train[["AccuracyUpper"]], digits = 2), ")", sep = "")
knn_byClass_train <- knn_conf_matrix_train$byClass
knn_sens_train <- knn_byClass_train["Sensitivity"]
knn_spec_train <- knn_byClass_train["Specificity"]
knn_PPV_train <- knn_byClass_train["Pos Pred Value"]
knn_NPV_train <- knn_byClass_train["Neg Pred Value"]
knn_prec_train <- knn_byClass_train["Precision"]
knn_rec_train <- knn_byClass_train["Recall"]
knn_F1_train <- knn_byClass_train["F1"]
knn_balAc_train <- knn_byClass_train["Balanced Accuracy"]

knn_overall_test <- knn_conf_matrix_test$overall
knn_accConf_test <- paste(round(knn_overall_test[["Accuracy"]], digits = 2), " (", round(knn_overall_test[["AccuracyLower"]], digits = 2), " - ", round(knn_overall_test[["AccuracyUpper"]], digits = 2), ")", sep = "")
knn_byClass_test <- knn_conf_matrix_test$byClass
knn_sens_test <- knn_byClass_test["Sensitivity"]
knn_spec_test <- knn_byClass_test["Specificity"]
knn_PPV_test <- knn_byClass_test["Pos Pred Value"]
knn_NPV_test <- knn_byClass_test["Neg Pred Value"]
knn_prec_test <- knn_byClass_test["Precision"]
knn_rec_test <- knn_byClass_test["Recall"]
knn_F1_test <- knn_byClass_test["F1"]
knn_balAc_test <- knn_byClass_test["Balanced Accuracy"]

KNN_train = c(knn_balAc_train, knn_sens_train, knn_spec_train, knn_PPV_train, knn_NPV_train, knn_prec_train, knn_rec_train, knn_F1_train)
KNN_test = c(knn_balAc_test, knn_sens_test, knn_spec_test, knn_PPV_test, knn_NPV_test, knn_prec_test, knn_rec_test, knn_F1_test)

rm(knn_sens_train, knn_spec_train, knn_PPV_train, knn_NPV_train, knn_prec_train, knn_rec_train, knn_F1_train, knn_balAc_train, knn_sens_test, knn_spec_test, knn_PPV_test, knn_NPV_test, knn_prec_test, knn_rec_test, knn_F1_test, knn_balAc_test)

# random forest custom
rf2_overall_train <- rf_conf_matrix_train2$overall
rf2_accConf_train <- paste(round(rf2_overall_train[["Accuracy"]], digits = 2), " (", round(rf2_overall_train[["AccuracyLower"]], digits = 2), " - ", round(rf2_overall_train[["AccuracyUpper"]], digits = 2), ")", sep = "")
rf2_byClass_train <- rf_conf_matrix_train2$byClass
rf2_sens_train <- rf2_byClass_train["Sensitivity"]
rf2_spec_train <- rf2_byClass_train["Specificity"]
rf2_PPV_train <- rf2_byClass_train["Pos Pred Value"]
rf2_NPV_train <- rf2_byClass_train["Neg Pred Value"]
rf2_prec_train <- rf2_byClass_train["Precision"]
rf2_rec_train <- rf2_byClass_train["Recall"]
rf2_F1_train <- rf2_byClass_train["F1"]
rf2_balAc_train <- rf2_byClass_train["Balanced Accuracy"]

rf2_overall_test <- rf_conf_matrix_test2$overall
rf2_accConf_test <- paste(round(rf2_overall_test[["Accuracy"]], digits = 2), " (", round(rf2_overall_test[["AccuracyLower"]], digits = 2), " - ", round(rf2_overall_test[["AccuracyUpper"]], digits = 2), ")", sep = "")
rf2_byClass_test <- rf_conf_matrix_test2$byClass
rf2_sens_test <- rf2_byClass_test["Sensitivity"]
rf2_spec_test <- rf2_byClass_test["Specificity"]
rf2_PPV_test <- rf2_byClass_test["Pos Pred Value"]
rf2_NPV_test <- rf2_byClass_test["Neg Pred Value"]
rf2_prec_test <- rf2_byClass_test["Precision"]
rf2_rec_test <- rf2_byClass_test["Recall"]
rf2_F1_test <- rf2_byClass_test["F1"]
rf2_balAc_test <- rf2_byClass_test["Balanced Accuracy"]

RF_train = c(rf2_balAc_train, rf2_sens_train, rf2_spec_train, rf2_PPV_train, rf2_NPV_train, rf2_prec_train, rf2_rec_train, rf2_F1_train)
RF_test = c(rf2_balAc_test, rf2_sens_test, rf2_spec_test, rf2_PPV_test, rf2_NPV_test, rf2_prec_test, rf2_rec_test, rf2_F1_test)

rm(rf2_sens_train, rf2_spec_train, rf2_PPV_train, rf2_NPV_train, rf2_prec_train, rf2_rec_train, rf2_F1_train, rf2_balAc_train, rf2_sens_test, rf2_spec_test, rf2_PPV_test, rf2_NPV_test, rf2_prec_test, rf2_rec_test, rf2_F1_test, rf2_balAc_test)

# AdaBoost
ada_overall_train <- ada_conf_matrix_train$overall
ada_accConf_train <- paste(round(ada_overall_train[["Accuracy"]], digits = 2), " (", round(ada_overall_train[["AccuracyLower"]], digits = 2), " - ", round(ada_overall_train[["AccuracyUpper"]], digits = 2), ")", sep = "")
ada_byClass_train <- ada_conf_matrix_train$byClass
ada_sens_train <- ada_byClass_train["Sensitivity"]
ada_spec_train <- ada_byClass_train["Specificity"]
ada_PPV_train <- ada_byClass_train["Pos Pred Value"]
ada_NPV_train <- ada_byClass_train["Neg Pred Value"]
ada_prec_train <- ada_byClass_train["Precision"]
ada_rec_train <- ada_byClass_train["Recall"]
ada_F1_train <- ada_byClass_train["F1"]
ada_balAc_train <- ada_byClass_train["Balanced Accuracy"]

ada_overall_test <- ada_conf_matrix_test$overall
ada_accConf_test <- paste(round(ada_overall_test[["Accuracy"]], digits = 2), " (", round(ada_overall_test[["AccuracyLower"]], digits = 2), " - ", round(ada_overall_test[["AccuracyUpper"]], digits = 2), ")", sep = "")
ada_byClass_test <- ada_conf_matrix_test$byClass
ada_sens_test <- ada_byClass_test["Sensitivity"]
ada_spec_test <- ada_byClass_test["Specificity"]
ada_PPV_test <- ada_byClass_test["Pos Pred Value"]
ada_NPV_test <- ada_byClass_test["Neg Pred Value"]
ada_prec_test <- ada_byClass_test["Precision"]
ada_rec_test <- ada_byClass_test["Recall"]
ada_F1_test <- ada_byClass_test["F1"]
ada_balAc_test <- ada_byClass_test["Balanced Accuracy"]

AB_train = c(ada_balAc_train, ada_sens_train, ada_spec_train, ada_PPV_train, ada_NPV_train, ada_prec_train, ada_rec_train, ada_F1_train)
AB_test = c(ada_balAc_test, ada_sens_test, ada_spec_test, ada_PPV_test, ada_NPV_test, ada_prec_test, ada_rec_test, ada_F1_test)

rm(ada_overall_train, ada_sens_train, ada_spec_train, ada_PPV_train, ada_NPV_train, ada_prec_train, ada_rec_train, ada_F1_train, ada_balAc_train, ada_overall_test, ada_sens_test, ada_spec_test, ada_PPV_test, ada_NPV_test, ada_prec_test, ada_rec_test, ada_F1_test, ada_balAc_test)

# assemble table
results_APAP_table <- round_df(data.frame(LR_train, LR_test, DT_train, DT_test, RF_train, RF_test, KNN_train, KNN_test, AB_train, AB_test), digits = 2)

# accuracy
accConf <- c(lr_accConf_train, lr_accConf_test, dt2_accConf_train, dt2_accConf_test, rf2_accConf_train, rf2_accConf_test, knn_accConf_train, knn_accConf_test, ada_accConf_train, ada_accConf_test)

results_APAP_table <- rbind(accConf, results_APAP_table)
rownames(results_APAP_table) <- c("Accuracy", "Balanced Accuracy", "Sensitivity", "Specificity", "PPV", "NPV", "Precision", "Recall", "F1 Score")

write_xlsx(results_APAP_table, "results_APAA_231101.xlsx")

pdf("results_APAP_231109.pdf", width = 18, height = 5)
results_APAP_table <- tableGrob(data.frame(results_APAP_table))
grid.arrange(results_APAP_table)
dev.off()

```
